Asymmetric relationships and their effects on coexistence: Supporting information

### 2023-02-01 paper/SupMat.Rnw
continued progress editing Methods and accompanying SI sections (mostly working on the SI)

%***DAN: Might need this text below, in the section on the log-normal fecundities lottery model, but not here
%Since we assumed log-normally distributed $B_i$, 
%$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
%Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
%to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
%and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into main text equation (\ref{MT-eq:storage_effects_gen}).

@@

We next decompose $\olr_{j \bs i}$ for the lottery model. We have 

@@

E_j = B_j = \exp(b_j)

@@

C_j=\frac{B_1 N_1 + B_2 N_2}{\delta N}
C_{j \bs i} = B_j/\delta = \exp(b_j)/\delta
r_j=\ln(1-\delta+E_j/C_j)

@@

&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(b_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta\frac{\exp(\sigma \tilde{b}_j + \mu_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(\sigma \tilde{b}_j - \sigma^2/2))]-\epsilon_j^0, \label{eq:decomp:contrib_of_var_E_j}

@@

where $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,

@@

&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(b_j)} \right) \right]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(\sigma \tilde{b}_j +\mu_j)} \right) \right]-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(-\sigma \tilde{b}_j+\sigma^2/2))]-\epsilon_j^0,\label{eq:decomp:contrib_of_var_C_j}

@@

where again $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,

@@

&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(b_j^\shp)}{\exp(b_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\sigma \tilde{b}_j^\shp + \mu_j)}{\exp(\sigma \tilde{b}_j + \mu_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))]-\olep_j^E-\olep_j^C-\epsilon_j^0, \label{eq:decomp:j:EshpC}

@@

$B_j^\shp$ is distributed in the same way as $B_j$ but is independent of it; $b_j^\shp$ is distributed in the same
way as $b_j$ but is independent of it; and $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and 
independent of each other. Continuing,

@@

&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:pECp}

@@

these results obtained from (\ref{eq:decomp:j:EC}) and (\ref{eq:decomp:j:EshpC}). Again, $\tilde{b}_j$ and 
$\tilde{b}_j^\shp$ are standard normally distributed and independent. Continuing,

@@

because we showed in the Theory section of the main text that $\olr_{j \bs i}^\nat = \olr_{j \bs i}$ due to the fact that

@@

&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:EChard}

@@

where again $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent of each other.
This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$ for the lottery model.

Using the above formulas, we observe that all the terms of the decomposition of $\olr_{i \bs i}$ depend on 
$\delta$, $\sigma$, and $\mu_i - \mu_j$, but not separately on $\mu_i$ or $\mu_j$. The terms of the decomposition
of $\olr_{j \bs i}$ depend on $\delta$ and $\sigma$, but do not depend on $\mu_i$ or $\mu_j$ at all.
As discussed in the main text and SI section \ref{SIsect:scaling_factors}, we consider two alternative
values for the scaling factors, $q_{ij}$, namely $q_{ij}=1$ and 
\begin{align}
q_{ij} &= \frac{\exp(\mu_i)}{(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)} \\
&= \frac{\exp(-\mu_i)\exp(\mu_i)}{\exp(-\mu_i)[(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)]} \\
&= \frac{1}{(1-\delta)\exp(-(\mu_i-\mu_j))+\delta}.\label{eq:decomp:qnot1}
\end{align}
These also either do not depend on $\mu_i$ or $\mu_j$ at all or they depend on the difference 
$\mu_i-\mu_j$. Thus each of the terms of the decomposition (\ref{MT-eq:full_decomp}) from the main text
depends only on $\mu_i-\mu_j$ and not separately on $\mu_i$ or $\mu_j$, as stated in the Methods section of the main text.

We now want to show that, for the lottery model, 
all elements of our decompositions (\ref{eq:decomp_overall_IGR}) and (\ref{eq:decomp_overall_rjbsi}), as well
as $q_{ij}$, and therefore all elements of our decomposition (\ref{eq:decomp_overall_delta}), are the same 
for our left- and right-tail associated cases, as was stated in Methods in the main text. Using the notation 
$(E_i^{(l)},C_{i \bs i}^{(l)})=(B_i^{(l)},B_j^{(l)}/\delta)=(\exp(b_i^{(l)}),\exp(b_j^{(l)})/\delta)$ and 
$(E_i^{(r)},C_{i \bs i}^{(r)})=(B_i^{(r)},B_j^{(r)}/\delta)=(\exp(b_i^{(r)}),\exp(b_j^{(r)})/\delta)$
for the left- and right-tail associated cases, respectively, we note that the marginal distributions of
$(b_i^{(l)},b_j^{(l)})$ are the same as those of $(b_i^{(r)},b_j^{(r)})$, and likewise the marginals of 
$(B_i^{(l)},B_j^{(l)}/\delta)$ are the same as those of $(B_i^{(r)},B_j^{(r)}/\delta)$.
So any component of (\ref{eq:decomp_overall_IGR}), or $q_{ij}$, will not depend on the choice
of our left- or right-tail associated noise cases as long as it depends only on marginal distributions.
But it is easy to see that $q_{ij}$ depends only on marginals, since $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$  
depends only on $\mu_i$, $\mu_j$ and $\delta$, and these are parameters of the marginal distributions of 
$(b_i,b_j)$. The alternative value $q_{ij}=1$ is constant, so also takes the same value for both our
left- and right-tail association noise cases. It is also easy to see from the definitions of $\epsilon_i^0$, 
$\olep_i^E$, $\olep_i^C$, and $\olep_i^{(E \shp C)}$ that these quantities only depend on marginals.

@@

We now examine $\olep_i^{[EC]}$, which, by (\ref{eq:decomp:contrib_ATA_1}), equals 
\begin{equation}
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
\end{equation}
But $(\ddot{b}_i,\ddot{b}_j)$ is a bivariate normal random variable constructed to have the same marginals and 
covariance as $(b_i,b_j)$, and the marginals and covariance of $(b_i^{(l)},b_j^{(l)})$ are the same as those of
$(b_i^{(r)},b_j^{(r)})$, so $(\ddot{b}_i^{(l)},\ddot{b}_j^{(l)})$ is distributed in the same way as 
$(\ddot{b}_i^{(r)},\ddot{b}_j^{(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
of the choice between our left- and right-tail-associated noise cases. Now examining 
$\E[\ln(1-\delta+\delta\exp(b_i-b_j))]$, we define $\beta_1=-b_2+\mu_1+\mu_2$ and $\beta_2=-b_1+\mu_1+\mu_2$,
and we note that $(\beta_1^{(l)},\beta_2^{(l)})$ is distributed in the same way as 
$(b_1^{(r)},b_2^{(r)})$; likewise, $(\beta_1^{(r)},\beta_2^{(r)})$ is distributed in the same way as 
$(b_1^{(l)},b_2^{(l)})$. But then we have 
\begin{align}
\E[\ln(1-\delta+\delta\exp(b_i^{(l)}-b_j^{(l)}))] &= \E[\ln(1-\delta+\delta\exp(-\beta_j^{(l)}+\mu_1+\mu_2-(-\beta_i^{(l)}+\mu_1+\mu_2)))] \\
&= \E[\ln(1-\delta+\delta\exp(\beta_i^{(l)}-\beta_j^{(l)}))] \\
&= \E[\ln(1-\delta+\delta\exp(b_i^{(r)}-b_j^{(r)}))], 
\end{align}
which completes the proof that $\olep_i^{[EC]}$ is independent of left- versus right-tail associated noise.
For the terms of the decomposition (\ref{eq:decomp_overall_rjbsi}), it suffices to look at the equations for those terms, 
above, in this section, and note that all those terms depend only on marginals.

\section{The $\delta=1$ case}\label{sect:delta1}

%The i terms
Setting $\delta=1$ in (\ref{eq:decomp:baseline}) gives 
\begin{equation}
\epsilon_i^0 = \mu_i - \mu_j.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E}) gives
\begin{align}
\olep_i^E &= \mu_i-\mu_j-\sigma^2/2-\epsilon_i^0 \\
&= -\sigma^2/2.
\end{align}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C}) gives
\begin{align}
\olep_i^C &= \mu_i - \mu_j + \sigma^2/2 -\epsilon_i^0 \\
&= \sigma^2/2.
\end{align}
Setting $\delta=1$ in (\ref{eq:decomp:last_term_primary}) gives
\begin{align}
\olep_i^{EC} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
&= 0.
\end{align}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) gives
\begin{align}
\olep_i^{(E \shp C)} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
&= 0.
\end{align}
Setting $\delta=1$ in (\ref{eq:decomp:storage_effects}) gives
\begin{align}
\olep_i^{(EC)} &= \mu_i - \mu_j - (\mu_i -\mu_j) \\
&= 0.
\end{align}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_ATA_2}) gives
\begin{equation}
\olep_i^{[E \nat C]}=0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_EC}) gives
\begin{equation}
\olep_i^{[EC]}=0.
\end{equation}
%The j terms
Setting $\delta=1$ in (\ref{eq:decomp:baseline_j}) gives
\begin{equation}
\epsilon_j^0 = 0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E_j}) gives
\begin{equation}
\olep_j^E = -\sigma^2/2.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C_j}) gives
\begin{equation}
\olep_j^C = \sigma^2/2.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:j:EC}) gives
\begin{equation}
\olep_j^{EC} = 0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:j:EshpC}) gives
\begin{equation}
\olep_j^{(E \shp C)} = 0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:j:pECp}) gives
\begin{equation}
\olep_j^{(EC)} = 0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:j:EnatC}) gives
\begin{equation}
\olep_j^{[E \nat C]} = 0.
\end{equation}
Setting $\delta=1$ in (\ref{eq:decomp:j:EChard}) gives
\begin{equation}
\olep_j^{[EC]}=0.
\end{equation}

For $\delta=1$, the second value of $q_{ij}$ we considered, namely
$q_{ij}=\frac{\exp(\mu_i)}{(1-\delta)\exp(\mu_j)+\delta \exp(\mu_i)}$, equals $1$, so both the
options we considered for $q_{ij}$ are $1$ in this case. Combining the above 
equations using $q_{ij}=1$ gives
\begin{align}
\Delta_i^0 &= \mu_i - \mu_j \\
\Delta_i^E &= 0  \\
\Delta_i^C &= 0  \\
\Delta_i^{(E \shp C)} &=  0 \\
\Delta_i^{[E \nat C]} &= 0  \\
\Delta_i^{[EC]} &=  0. 
\end{align}
Thus the only contribution to coexistence for the $\delta=1$ case is $\Delta_i^0 = \mu_i - \mu_j$,
which does not depend on $\sigma$ and which is non-positive whenever $\mu_i \leq \mu_j$ (which we assumed,
without loss of generality).
So there is no robust coexistence in the $\delta=1$ case for our simple model.

@@

We here elaborate how the terms in the decompositions (\ref{eq:decomp_overall_IGR}), (\ref{eq:decomp_overall_rjbsi})
and (\ref{eq:decomp_overall_delta}) are computed rapidly, for the lottery model.

First, for a large integer $M$ (we used $M=\Sexpr{M}$),
carry out the following steps once.
\begin{enumerate}
\item Generate $M$ independent points $u^{(k)}$, $k=1,\ldots,M$, from a standard normal distribution.
\item Generate $M$ independent points $(\tilde{b}_i^{(k)},\tilde{b}_j^{(k)})$, as in SI section \ref{sect:noise}, which are left-tail associated
and which have standard-normal marginals. 
\item Again as in SI section \ref{sect:noise}, generate $M$ independent points $(\tilde{\ddot{b}}_i^{(k)},\tilde{\ddot{b}}_j^{(k)})$, 
$k=1,\ldots,M$, from a bivariate normal distribution with component variances $1$ and Pearson correlation the same as 
$(\tilde{b}_i^{(k)},\tilde{b}_j^{(k)})$.
\end{enumerate}
These data are stored and used repeatedly below. This provides computational efficiency because 
random number generation is computationally expensive.

>>>>

@@

We next decompose $\olr_{j \bs i}$ for the log-normal fecundities lottery model [see (\ref{eq:decomp_overall_rjbsi})]. We have 

@@

E_j = B_j = \exp(\sigma b_j + \mu_j)

@@

C_{j \bs i} = B_j/\delta = \exp(\sigma b_j + \mu_j)/\delta

@@

&= \E \left[ \ln \left(1-\delta+\delta\frac{\exp(\sigma b_j + \mu_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(\sigma b_j - \sigma^2/2))]-\epsilon_j^0, \label{eq:decomp:contrib_of_var_E_j}

@@

@@

&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(\sigma b_j +\mu_j)} \right) \right]-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(-\sigma b_j+\sigma^2/2))]-\epsilon_j^0,\label{eq:decomp:contrib_of_var_C_j}

@@

where again $b_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,

@@

&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\sigma b_j^\shp + \mu_j)}{\exp(\sigma b_j + \mu_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(\sigma(b_j^\shp - b_j)))]-\olep_j^E-\olep_j^C-\epsilon_j^0, \label{eq:decomp:j:EshpC}

@@

$B_j^\shp$ is distributed in the same way as $B_j$ but is independent of it; and $b_j^\shp$ is distributed in the same
way as $b_j$ but is independent of it. Continuing,

@@

&= -\E[\ln(1-\delta+\delta \exp(\sigma (b_j^\shp - b_j)))].\label{eq:decomp:j:pECp}

@@

@@

because we showed in SI section \ref{SIsect:TheoryForLotteryBoth} that $\olr_{j \bs i}^\nat = \olr_{j \bs i}$, 
due to the fact that

@@

&= - \E[\ln(1-\delta+\delta \exp(\sigma(b_j^\shp - b_j)))],\label{eq:decomp:j:EChard}

@@

where again $b_j$ and $b_j^\shp$ are standard normally distributed and independent of each other.
This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$ for the log-normal
fecundities lottery model.

Using the above formulas, we observe that all the terms of the decomposition of $\olr_{i \bs i}$ depend on 
$\delta$, $\sigma$, and $\mu_i - \mu_j$, but not separately on $\mu_i$ or $\mu_j$. The terms of the decomposition
of $\olr_{j \bs i}$ depend on $\delta$ and $\sigma$, but do not depend on $\mu_i$ or $\mu_j$ at all.
As discussed in the main text and SI section \ref{SIsect:scaling_factors}, we consider two alternative
values for the scaling factors, $q_{ij}$, namely $q_{ij}=1$ and a value which is shown in 
SI section \ref{SIsect:TheoryForLotteryBoth}
to equal, for the lottery model (both versions), the quantity
\begin{equation}
q_{ij} = \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta}.
\end{equation}
For the log-normal fecundities lottery model, this is
\begin{align}
q_{ij} &= \frac{\exp(\mu_i)}{(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)} \\
&= \frac{\exp(-\mu_i)\exp(\mu_i)}{\exp(-\mu_i)[(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)]} \\
&= \frac{1}{(1-\delta)\exp(-(\mu_i-\mu_j))+\delta}.\label{eq:decomp:qnot1}
\end{align}
Thus the $q_{ij}$ either do not depend on $\mu_i$ or $\mu_j$ at all or they depend on the difference 
$\mu_i-\mu_j$, rather than depending separately on $\mu_i$ or $\mu_j$. Thus each of the terms of the 
decomposition (\ref{MT-eq:full_decomp}) from the main text
depends only on $\mu_i-\mu_j$ and not separately on $\mu_i$ or $\mu_j$, as stated in the Methods section of the main text.

We now want to show that, for the log-normal fecundities lottery model, 
all elements of our decompositions (\ref{eq:decomp_overall_IGR}) and (\ref{eq:decomp_overall_rjbsi}), as well
as $q_{ij}$, and therefore all elements of our decomposition (\ref{eq:decomp_overall_delta}), are the same 
for our left- and right-tail associated cases, as was stated in Methods in the main text. Using the notation 
$(E_i^{(l)},C_{i \bs i}^{(l)})=(B_i^{(l)},B_j^{(l)}/\delta)=(\exp(\sigma b_i^{(l)} + \mu_i),
\exp(\sigma b_j^{(l)} + \mu_j)/\delta)$ and 
$(E_i^{(r)},C_{i \bs i}^{(r)})=(B_i^{(r)},B_j^{(r)}/\delta)=(\exp(\sigma b_i^{(r)}+\mu_i),\exp(\sigma b_j^{(r)}+\mu_j)/\delta)$
for the left- and right-tail associated cases, respectively, we note that the marginal distributions of
$(b_i^{(l)},b_j^{(l)})$ are the same as those of $(b_i^{(r)},b_j^{(r)})$, and so likewise the marginals of 
$(B_i^{(l)},B_j^{(l)}/\delta)$ are the same as those of $(B_i^{(r)},B_j^{(r)}/\delta)$.
So any component of (\ref{eq:decomp_overall_IGR}), or $q_{ij}$, will not depend on the choice
of our left- or right-tail associated noise cases as long as it depends only on marginal distributions.
But it is easy to see that $q_{ij}$ depends only on marginals, since $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$  
depends only on $\mu_i$, $\mu_j$ and $\delta$, and these are parameters of the marginal distributions of 
$(b_i,b_j)$. The alternative value $q_{ij}=1$ is constant, so also takes the same value for both our
left- and right-tail association noise cases. It is also easy to see from the definitions of $\epsilon_i^0$, 
$\olep_i^E$, $\olep_i^C$, and $\olep_i^{(E \shp C)}$ that these quantities only depend on marginals.

@@

We now examine $\olep_i^{[EC]}$, which, by (\ref{eq:decomp:contrib_ATA_1}), equals 
\begin{equation}
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(b_i-b_j)+\mu_i-\mu_j)]-
\E[\ln(1-\delta+\delta\exp(\sigma(b_i^\nat-b_j^\nat)+\mu_i-\mu_j)].\label{eq:rep1}
\end{equation}
But $(b_i^\nat,b_j^\nat)$ is a bivariate normal random variable constructed to have the same marginals and 
covariance as $(b_i,b_j)$, and the marginals and covariance of $(b_i^{(l)},b_j^{(l)})$ are the same as those of
$(b_i^{(r)},b_j^{(r)})$, so $(b_i^{\nat,(l)},b_j^{\nat,(l)})$ is distributed in the same way as 
$(b_i^{\nat,(r)},b_j^{\nat,(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
of the choice between our left- and right-tail-associated noise cases. Now examining 
$\E[\ln(1-\delta+\delta\exp(\sigma(b_i-b_j)+\mu_i-\mu_j))]$, 
we define $\beta_1=-b_2$ and $\beta_2=-b_1$,
and we note that $(\beta_1^{(l)},\beta_2^{(l)})$ is distributed in the same way as 
$(b_1^{(r)},b_2^{(r)})$; likewise, $(\beta_1^{(r)},\beta_2^{(r)})$ is distributed in the same way as 
$(b_1^{(l)},b_2^{(l)})$. So we then have 
\begin{align}
\E[\ln(1-\delta+\delta\exp(\sigma(b_i^{(l)}-b_j^{(l)})+\mu_i-\mu_j))] 
&= \E[\ln(1-\delta+\delta\exp(\sigma(\beta_i^{(r)}-\beta_j^{(r)})+\mu_i+\mu_j))] \\
&= \E[\ln(1-\delta+\delta\exp(\sigma(-b_j^{(r)}+b_i^{(r)})+\mu_i-\mu_j))] \\
&= \E[\ln(1-\delta+\delta\exp(\sigma(b_i^{(r)}-b_j^{(r)})+\mu_i-\mu_j))], 
\end{align}
which completes the proof that $\olep_i^{[EC]}$ is independent of left- versus right-tail associated noise.
For the terms of the decomposition (\ref{eq:decomp_overall_rjbsi}), it suffices to look at the equations for those terms, 
above, in this section, and note that all those terms depend only on marginals.

\subsection{Theory and plotting simplifications for the beta fecundities lottery model}\label{SIsect:TheoryForLotteryBeta}
%***DAN: Fill this in.

%***DAN: I am 90% sure that once I re-read the comment of Ellner saying we do not need to consider really
%big \delta, we will be able to delete the below.
% \section{The $\delta=1$ case}\label{sect:delta1}
% 
% %The i terms
% Setting $\delta=1$ in (\ref{eq:decomp:baseline}) gives 
% \begin{equation}
% \epsilon_i^0 = \mu_i - \mu_j.
% \end{equation}
% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E}) gives
% \begin{align}
% \olep_i^E &= \mu_i-\mu_j-\sigma^2/2-\epsilon_i^0 \\
% &= -\sigma^2/2.
% \end{align}
% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C}) gives
% \begin{align}
% \olep_i^C &= \mu_i - \mu_j + \sigma^2/2 -\epsilon_i^0 \\
% &= \sigma^2/2.
% \end{align}
% Setting $\delta=1$ in (\ref{eq:decomp:last_term_primary}) gives
% \begin{align}
% \olep_i^{EC} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
% &= 0.
% \end{align}
% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) gives
% \begin{align}
% \olep_i^{(E \shp C)} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
% &= 0.
% \end{align}
% Setting $\delta=1$ in (\ref{eq:decomp:storage_effects}) gives
% \begin{align}
% \olep_i^{(EC)} &= \mu_i - \mu_j - (\mu_i -\mu_j) \\
% &= 0.
% \end{align}
% Setting $\delta=1$ in (\ref{eq:decomp:contrib_ATA_2}) gives
% \begin{equation}
% \olep_i^{[E \nat C]}=0.
% \end{equation}
% Setting $\delta=1$ in (\ref{eq:decomp:contrib_EC}) gives
% \begin{equation}
% \olep_i^{[EC]}=0.
% \end{equation}
% %The j terms
% Setting $\delta=1$ in (\ref{eq:decomp:baseline_j}) gives
% \begin{equation}
% \epsilon_j^0 = 0.
% \end{equation}
% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E_j}) gives
% \begin{equation}
% \olep_j^E = -\sigma^2/2.
% \end{equation}
% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C_j}) gives
% \begin{equation}
% \olep_j^C = \sigma^2/2.
% \end{equation}
% Setting $\delta=1$ in (\ref{eq:decomp:j:EC}) gives
% \begin{equation}
% \olep_j^{EC} = 0.
% \end{equation}
% Setting $\delta=1$ in (\ref{eq:decomp:j:EshpC}) gives
% \begin{equation}
% \olep_j^{(E \shp C)} = 0.
% \end{equation}
% Setting $\delta=1$ in (\ref{eq:decomp:j:pECp}) gives
% \begin{equation}
% \olep_j^{(EC)} = 0.
% \end{equation}
% Setting $\delta=1$ in (\ref{eq:decomp:j:EnatC}) gives
% \begin{equation}
% \olep_j^{[E \nat C]} = 0.
% \end{equation}
% Setting $\delta=1$ in (\ref{eq:decomp:j:EChard}) gives
% \begin{equation}
% \olep_j^{[EC]}=0.
% \end{equation}
% 
% For $\delta=1$, the second value of $q_{ij}$ we considered, namely
% $q_{ij}=\frac{\exp(\mu_i)}{(1-\delta)\exp(\mu_j)+\delta \exp(\mu_i)}$, equals $1$, so both the
% options we considered for $q_{ij}$ are $1$ in this case. Combining the above 
% equations using $q_{ij}=1$ gives
% \begin{align}
% \Delta_i^0 &= \mu_i - \mu_j \\
% \Delta_i^E &= 0  \\
% \Delta_i^C &= 0  \\
% \Delta_i^{(E \shp C)} &=  0 \\
% \Delta_i^{[E \nat C]} &= 0  \\
% \Delta_i^{[EC]} &=  0. 
% \end{align}
% Thus the only contribution to coexistence for the $\delta=1$ case is $\Delta_i^0 = \mu_i - \mu_j$,
% which does not depend on $\sigma$ and which is non-positive whenever $\mu_i \leq \mu_j$ (which we assumed,
% without loss of generality).
% So there is no robust coexistence in the $\delta=1$ case for our simple model.

@@

\subsection{Efficient computation for the log-normal fecundities lottery model}\label{sect:efficient_computation_lognormal}

We here elaborate how the terms in the decompositions (\ref{eq:decomp_overall_IGR}), (\ref{eq:decomp_overall_rjbsi})
and (\ref{eq:decomp_overall_delta}) are computed rapidly, for the log-normal fecundities lottery model.

First, for a large integer $M$ (we used $M=\Sexpr{M}$),
carry out the following steps once.
\begin{enumerate}
\item Generate $M$ independent points $u^{(k)}$, $k=1,\ldots,M$, from a standard normal distribution.
\item Generate $M$ independent points $(b_i^{(k)},b_j^{(k)})$, $k=1,\ldots,M$, as in SI section \ref{sect:noise}, 
which are left-tail associated and which have standard-normal marginals. 
\item Again as in SI section \ref{sect:noise}, generate $M$ independent points $(b_i^{\nat,(k)},b_j^{\nat,(k)})$, 
$k=1,\ldots,M$, from a bivariate normal distribution with component variances $1$ and Pearson correlation the same as 
$(b_i^{(k)},b_j^{(k)})$.
\end{enumerate}
These data are stored and used repeatedly below. This provides computational efficiency because 
random number generation is computationally expensive.

Next, given values of the parameters $\delta$, $\mu_i - \mu_j$, and $\sigma$, proceed as follows to get estimates of the 
terms of the decomposition (\ref{eq:decomp_overall_IGR}).
\begin{enumerate}
\item Compute $\epsilon_i^0$ following (\ref{eq:decomp:baseline}), which is a simple formula of $\delta$ and 
$\mu_i - \mu_j$, not requiring use of the random samples described above. This gives an exact value of
$\epsilon_i^0$, not an estimate.
\item Estimate $\olep_i^E$, following (\ref{eq:decomp:contrib_of_var_E}), as 
$\widehat{\olep_i^E}=\mean_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0$, an estimate with 
standard error $\se \left( \widehat{\olep_i^E} \right) = \sd_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]/\sqrt{M}$,
where $\sd(\cdot)$ denotes sample standard deviation.
\item Estimate $\olep_i^C$, following (\ref{eq:decomp:contrib_of_var_C}), as 
$\widehat{\olep_i^C}=\mean_k [\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-\epsilon_i^0$, which has 
standard error $\se \left( \widehat{\olep_i^C} \right) = \sd_k[\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_i^\shp-\tilde{b}_j^\shp$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) is normally distributed
with mean $0$ and variance $2$, so we can estimate $\olep_i^{(E \shp C)}$, following (\ref{eq:decomp:contrib_of_EC_var_per_se}),
as $\widehat{\olep_i^{(E \shp C)}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0.$ The standard error of this expression is 
$\se \left( \widehat{\olep_i^{(E \shp C)}} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma \sqrt{2} u^{(k)}+\mu_i-\mu_j))-
\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$. 
\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_EC}), as
$\widehat{\olep_i^{[E \nat C]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$. The standard error of this expression is 
$\se \left( \widehat{\olep_i^{[E \nat C]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))]/\sqrt{M}$
and $s_2=\sd_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]/\sqrt{M}$.
\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_ATA_1}), as 
$\widehat{\olep_i^{[EC]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))]-
\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))].$ The standard
error of this expression is $\se \left( \widehat{\olep_i^{[EC]}} \right)=\sqrt{s_3^2+s_1^2}$, where
$s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}.$
\item Finally, estimate $\olr_{i \bs i}$, following (\ref{eq:base_IGR}), as 
$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))],$
which has standard error $s_3$ from above.
\end{enumerate}
It is easy to see that $\widehat{\olr_{i \bs i}}=\epsilon_i^0+\widehat{\olep_i^E}+\widehat{\olep_i^C}+\widehat{\olep_i^{(E \shp C)}}+
\widehat{\olep_i^{[E \nat C]}}+\widehat{\olep_i^{[EC]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
at the level of the estimators, as well
